---
title: "Idaho Shape Complexity Analysis"
author: "Darin Schulte"
date: "September 9, 2016"
output:
  html_document: 
    toc: true
    toc_depth: 6
---


###Load Necessary Packages      
 
```{r}
# Spatial data and analysis packages
library(rgdal)
getGDALVersionInfo()
library(gdalUtils)
library(tools)
library(raster)
library(sp)
library(shapefiles)
library(reshape2)
library(spatstat)
library(maptools)
library(cleangeo)
library(rgeos)

# Plotting packages
library(ggplot2)
library(grid)
library(Rmisc)
library(GGally)

# General packages for data manipulation and statistics
library(data.table)
library(dtplyr)
library(stats)
library(plyr)
library(car)
library(lme4)
library(nlme)
library(MASS)
library(bbmle)
```   

<br/>   

-----------    


## General notes   
### Plot descriptions   
Plots **Q1 - Q10** and **Q19 - Q26** were not grazed during the data collection period. These are the focus of this analysis. 

## Set the working directory   
```{r}
# Set working directory
setwd("C:\\Users\\darin\\Dropbox\\wd_R\\idaho_analysis\\working_data")
getwd()
```      


## Create list of shapefiles   
List of shapefiles contains only the **cover** shapefiles (i.e., the polygon files not the point files) and only those plots that were not grazed during the study period.   

```{r, eval = FALSE}
# Create list of file names (just the .shp file names)
shpFiles <- list.files("./msData/shapefiles", pattern = glob2rx("*.shp"), full.names = TRUE)
head(shpFiles)

# Extract base names (i.e., file names without full paths)
baseNames <- basename(shpFiles)
head(baseNames)

# Remove .shp extension to get just the plot name and year
baseNames <- file_path_sans_ext(baseNames)
#head(baseNames)
# Split file names to get access to the plot name and type (cover vs density)
plotNames <- strsplit(baseNames, '_', '\\.')
length(plotNames)
head(plotNames)

plotsList <- list(NA, rep(length(plotNames)))
for(i in 1:length(plotNames)) {
    plotsList[i] <- plotNames[[i]][1]
}
plotsList <- unlist(plotsList)

# Get subset of list containing only non-grazed plots (Q1 - Q10 & Q19 - Q26)
u_g_plots <- c("Q1", "Q2", "Q3", "Q4", "Q5", "Q6", "Q7", "Q8", "Q9", "Q10", "Q19", "Q20", "Q21", "Q22", "Q23", "Q24", "Q25", "Q26")
testList <-  plotsList %in% u_g_plots
testList
plotsList_ng <- shpFiles[testList==TRUE]
plotsList_ng

# Get a sublist of only cover (polygons) shapefiles (the density shapefiles are points not polygons)
baseNames <- basename(plotsList_ng)
baseNames <- file_path_sans_ext(baseNames)
plotNames <- strsplit(baseNames, "_", "\\.")
plotsList <- list(NA, rep(length(plotNames)))
for (i in 1:length(plotNames)) {
    plotsList[i] <- plotNames[[i]][3]
}
plotsList <- unlist(plotsList)
u_g_type <- "C"
testList <- plotsList %in% u_g_type
testList
plotsList_ng_C <- plotsList_ng[testList==TRUE]
plotsNG <- plotsList_ng_C
# Plot names used below in reassigning cleaned geometries to shapefiles
cleanedNames <- basename(plotsNG)
cleanedNames <- file_path_sans_ext(cleanedNames)
cleanedNames <- paste("./msData/cleanedShapeFiles/", cleanedNames, "_cleaned", sep="")

# Remove temp variables
#rm(plotsList_ng_C, plotsList_ng)
```      

<br/>   
 
---------------------   

## Shape index processing   

For each shapefile:   

* Clean the geometries for all polygons    
* Exclude the polygons that touch the bounding box for that plot    
* Calculate the Shape Complexity Index for each polygon (non-square representation)   
* Append results to dataframe     


### Cleaning the geometries   
**NOTE**: Q2_36_C_cleaned required manual editing of one of the polygons in QGIS. After editing, the shapefile resulting from the following code chunk was replaced with the manually edited version.   
```{r}
# Cleaning the geometries
######################################
# NOTE: This only needs to be run once
######################################
#dir.create("./msData/cleanedShapeFiles")

# Define a function to clean the geometries of the shapefiles and write them out to new shapefiles
cleanGeoms <- function(polygonFileList) {
    for(i in 1:length(polygonFileList)) {
        polyFile <- readShapePoly(polygonFileList[[i]])
        # Inspect the spatial objects to detect potential geometry issues and make a summary
        #   source: https://cran.r-project.org/web/packages/cleangeo/vignettes/quickstart.html
        report <- clgeo_CollectionReport(polyFile)

        # Clean the spatial objects 
        polyFile.clean <- clgeo_Clean(polyFile, strategy = "BUFFER") #Default strategy "POLYGONIZE" errors out

        polyFile.clean@data <- polyFile.clean@data[, c(1, 2)] #Only need "2", but spdf won't allow you to only choose one column
        # Write out to a new shapefile (that represents the cleaned geometry)
        writeSpatialShape(polyFile.clean, cleanedNames[i])
        }
    }
#length(cleanedNames)
cleanGeoms(plotsNG)
```   

### Exclude polygons touching bounding box   
```{r}
######################################
# NOTE: This only needs to be run once
######################################
#dir.create("./msData/nonEdgePolys")

# Create list of file names of 'cleaned' polygons(just the .shp file names)
clnShpFiles <- list.files("./msData/cleanedShapeFiles", pattern = glob2rx("*.shp"), full.names = TRUE)

# Define a function to test for intersection with bounding box
removeEdgePolys <- function(shapeFile) {
    shpToSPDF <- readShapePoly(shapeFile)
    
    # Create SpatialPolygons object out of the bounding box of the plot
    bboxPolygon <- as(extent(shpToSPDF), 'SpatialPolygons')

    # Test for intersection with bounding box 
    edgePolysTest <- gContainsProperly(bboxPolygon, shpToSPDF, byid = TRUE)
    
    # Get only those polygons that do not intersect bbox and save to shapefile
    spdf_edgeRmv <- shpToSPDF[c(edgePolysTest), ]
    shpBaseName <- basename(shapeFile)
    shpBaseName <- file_path_sans_ext(shpBaseName)
    edgeRmvNames <- paste("./msData/nonEdgePolys/", shpBaseName, "_edgeRmv", sep="")
    writeSpatialShape(spdf_edgeRmv, edgeRmvNames)
}

# Run function on all plots
lapply(clnShpFiles, function(x) {removeEdgePolys(x)})
```   

### Calculating Shape Complexity   
```{r}

#################################################################
#TODO: Make edits so that necessary variables are retained for later processes 
# even though the processing code chunks that generate the new shapefiles will not be run
####################################################################

#dir.create("./msData/sciPolys")

# Create list of file names of 'non-edge' polygons (created in previous code chunk)
noEdgeShpFiles <- list.files("./msData/nonEdgePolys", pattern = glob2rx("*.shp"), full.names = TRUE)

# Define function to calculate area, perimeter and shape complexity index for all polygons
shpCompInd <- function(shapes) {
    shpToSPDF <- readShapePoly(shapes)
    shpBaseName <- basename(shapes)
    shpBaseName <- file_path_sans_ext(shpBaseName)
    sciNames <- paste("./msData/sciPolys/", shpBaseName, "_sci", sep="")
    # Keep only ID and species name @data fields
    shpToSPDF@data <- shpToSPDF@data[, c(1, 4)]
    
    # Calculate area of each polygon and append 'area' column to @data
    shpToSPDF$area <- gArea(shpToSPDF, byid = TRUE)
    # Calculate perimeter of each polygon and append 'perimeter' column to @data
    shpToSPDF$perimeter <- gLength(shpToSPDF, byid = TRUE)
    
    # Calculate shape complexity index and append 'sci' column to @data
    shpToSPDF$sci <- shpToSPDF$perimeter/(2*pi*(sqrt(shpToSPDF$area/pi)))
    
    # Write out to shapefile overwriting previous file
    writeSpatialShape(shpToSPDF, sciNames)
}

# Calculate shape complexity for all polygons in list of spdfs
lapply(noEdgeShpFiles, function(x) {shpCompInd(x)})


# T-test 
t.test(shpToSPDF$sci, circles)

```   