---
title: "Shortgrass Steppe Shape Complexity Analysis"
author: "Darin Schulte"
date: "October 2, 2016"
output:
  html_document: 
    toc: true
    toc_depth: 6
---


###Load Necessary Packages      
 
```{r}
# Spatial data and analysis packages
library(rgdal)
getGDALVersionInfo()
library(gdalUtils)
library(tools)
library(raster)
library(sp)
library(shapefiles)
library(reshape2)
library(spatstat)
library(maptools)
library(cleangeo)
library(rgeos)

# Plotting packages
library(ggplot2)
library(grid)
library(Rmisc)
library(GGally)

# General packages for data manipulation and statistics
library(data.table)
library(dtplyr)
library(dplyr)
library(stats)
library(plyr)
library(car)
library(lme4)
library(nlme)
library(MASS)
library(bbmle)
```   

<br/>   

-----------    


## General notes   

### This rmd file is in the idaho data folder. The working directory cannot be changed as a result (even though this script deals with sgs data).   


## Set the working directory   
```{r}
# Set working directory
setwd("C:\\Users\\darin\\Dropbox\\wd_R\\idaho_analysis\\working_data")
getwd()
```      


## Create list of shapefiles   
List of shapefiles contains only the **cover** shapefiles (i.e., the polygon files not the point files).   
```{r, eval = FALSE}
# Create list of file names (just the .shp file names)
sgsDirs <- list.dirs("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/shapefiles")
# First item is the parent folder. Exclude that from shapefile directories list
sgsDirs <- sgsDirs[-1]

# Get paths to all shapefiles by iterating through the list of directories and extracting all shapefiles
shpFiles <- lapply(sgsDirs, function(x) {list.files(x, pattern = glob2rx("*.shp"), full.names = TRUE)})
# Get just the poly files (not the pnt files)
polyFiles <- lapply(shpFiles, function(x) {grep(x, pattern = "poly_")})
polysList <- lapply(1: length(shpFiles), function(x) {
    shpFiles[x] <- shpFiles[[x]][polyFiles[[x]]]
})
# Exclude all *_24 plots due to inconsistent mapping of B. gracilis and B. dactyloides (see Chu et al 2009 for reason)
polysList <- polysList[-c(3, 9, 15, 21)]

# Extract base names (i.e., file names without full paths)
baseNames <- lapply(polysList, function(x) {basename(x)})
head(baseNames)

# Remove .shp extension to get just the plot name and year
baseNames <- lapply(baseNames, function(x) {file_path_sans_ext(x)})


# Plot names used below in reassigning cleaned geometries to shapefiles
cleanedNames <- lapply(baseNames, function(x) {paste("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/cleanedShapeFiles/", x, "_cleaned", sep="")})
```      

<br/>   
 
---------------------   

## Shape index processing   

For each shapefile:   

* Clean the geometries for all polygons    
* Exclude the polygons that touch the bounding box for that plot    
* Calculate the Shape Complexity Index for each polygon    
* Append results to dataframe     


### Cleaning the geometries   
**NOTE**: Q2_36_C_cleaned required manual editing of one of the polygons in QGIS. After editing, the shapefile resulting from the following code chunk was replaced with the manually edited version.   
```{r}
# Cleaning the geometries
######################################
# NOTE: This only needs to be run once
######################################
# dir.create("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/cleanedShapeFiles")

# Sometimes polygons have multiple assignments for Species. Keep all as without this, there are often very few B. grama. All multi-species
#    assigments have BOGR as one of them
keepSpecies <- c("Bouteloua gracilis", "BOGR/CAEL", "BOGR/BUDA", "BOGR/SIHY")

# Define function to iterate through all shapefiles and clean goemetries. Write results to new shapefiles
cleanGeoms <- lapply(1:length(polysList), function(x) {
    lapply(1:length(polysList[[x]]), function(i) {
        polyTemp <- readShapePoly(polysList[[x]][i])
        # Get just Blue grama and multi-species assignments with BOGR as one 
        polyTemp <- polyTemp[polyTemp@data$Species %in% keepSpecies, ]
        #polyTemp <- polyTemp[b_grama, ]
        # Inspect
        report <- clgeo_CollectionReport(polyTemp)
        # Clean
        polyTemp.clean <- clgeo_Clean(polyTemp, strategy = "BUFFER") 
        # Get @data for SP_ID, SP_ID_1, and Species
        polyTemp.clean@data <- polyTemp.clean@data[, 1:3] 
        # Write out new shapefile representing cleaned geometries for B. gracilis only
        writeSpatialShape(polyTemp.clean, cleanedNames[[x]][i])
    })
})
# rm(cleanGeoms)
```   

### Exclude polygons touching bounding box   
```{r}
######################################
# NOTE: This only needs to be run once
######################################
#dir.create("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/nonEdgePolys")

# Create list of file names of 'cleaned' polygons(just the .shp file names)
clnShpFiles <- list.files("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/cleanedShapeFiles", pattern = glob2rx("*.shp"), full.names = TRUE)

# Iterate through all shapefiles to test for intersection with bounding box
lapply(clnShpFiles, function(x) {
    shpToSPDF <- readShapePoly(x)
    # Create SpatialPolygons object out of bounding box for the plot
    bboxPolygon <- as(extent(shpToSPDF), "SpatialPolygons")
    # Test for intersection with bounding box
    edgePolysTest <- gContainsProperly(bboxPolygon, shpToSPDF, byid = TRUE)
    # Get only those polygons that do not intersect the bbox and save to shapefile
    spdf_edgeRmv <- shpToSPDF[c(edgePolysTest), ]
    # Correct for extra @data columns being generated (calling writeShapePoly still generates an extra SP_ID column below)
    spdf_edgeRmv@data <- spdf_edgeRmv@data[, 2:4]
    colnames(spdf_edgeRmv@data) <- c("SP_ID", "SP_ID_1", "Species")
    # Create new file names
    shpBaseName <- basename(x)
    shpBaseName <- file_path_sans_ext(shpBaseName)
    edgeRmvNames <- paste("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/nonEdgePolys/", shpBaseName, "_edgeRmv", sep="")
    # Write to shapefiles
    writeSpatialShape(spdf_edgeRmv, edgeRmvNames)
})
```   

### Calculating Shape Complexity     
NOTE: This step only needs to be run once. Shapefiles are created for each plot and contain the shape complexity index in their respective data fields.   
```{r}
# dir.create("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sciPolys")

# Create list of file names of 'non-edge' polygons (created in previous code chunk)
noEdgeShpFiles <- list.files("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/nonEdgePolys", pattern = glob2rx("*.shp"), full.names = TRUE)

# Iterate through all shapefiles and calculate area, perimeter and shape complexity index for all polygons
lapply(noEdgeShpFiles, function(x) {
    shpToSPDF <- readShapePoly(x)
    shpBaseName <- basename(x)
    shpBaseName <- file_path_sans_ext(shpBaseName)
    sciNames <- paste("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sciPolys/", shpBaseName, "_sci", sep="")
    # Keep only ID and Species name from @data fields
    shpToSPDF@data <- shpToSPDF@data[, 2:4]
    colnames(shpToSPDF@data) <- c("SP_ID", "SP_ID_1", "Species")
    
    # Calculate area of each polygon and append 'area' column to @data
    shpToSPDF$area <- gArea(shpToSPDF, byid = TRUE)
    # Calculate perimeter of each polygon and append 'perimeter' column to @data
    shpToSPDF$perimeter <- gLength(shpToSPDF, byid = TRUE)
    
    # Calculate shape complexity index and append 'sci' column to @data
    shpToSPDF$sci <- shpToSPDF$perimeter/(2*pi*(sqrt(shpToSPDF$area/pi)))
    
    # Write out to shapefile
    writeSpatialShape(shpToSPDF, sciNames)
})
```   

### Gather shape complexity index values from all shapefiles   
```{r}
# Create list of file names for shapefiles with 'sci' fields
sciPolys <- list.files("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sciPolys", pattern = glob2rx("*.shp"), full.names=TRUE)

# Empty dataframe 
sciDF <- data.frame(matrix(ncol = 10, nrow = 30000))
colnames(sciDF) <- c("treatment", "plot", "year", "SP_ID", "SP_ID_1", "SP_ID_2", "Species", "area", "perimeter", "sci")

# Iterate through all shapefiles and store 'sci' values in a dataframe
sciDF <- do.call("rbind", lapply(sciPolys, function(x) {
    shpToSPDF <- readShapePoly(x)
    sciData <- shpToSPDF@data
    shpBaseName <- basename(x)
    shpBaseName <- file_path_sans_ext(shpBaseName)
    split <- strsplit(shpBaseName, "poly_")
    split <- strsplit(split[[1]][2], "[_]")
    split <- do.call(rbind.data.frame, split)
    split <- split[,1:3]
    sciTemp <- cbind(split, sciData)
    colnames(sciTemp) <- c("treatment", "plot", "year", "SP_ID", "SP_ID_1", "SP_ID_2", "Species", "area", "perimeter", "sci")
    return(sciTemp)
}))
# # Write results to files
# saveRDS(sciDF, file="C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_shapeComplexity.rds")
# write.csv(sciDF, file="C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_shapeComplexity.csv")
```      

### Shape complexity analysis   
```{r}
# Read in Shape Complexity data
sci_DF <- readRDS("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_shapeComplexity.rds")

# Read in weather data
weather <- read.csv("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/daily_climate_data.csv")
head(weather)
# Melt to look at annual precip
# Melt sci_DF for analysis
sciMelt <- melt(sci_DF, id = c("treatment", "plot", "year"), measure.vars = "sci")
# Subset sciMelt to exclude the arbitrary square polygons
sciNoSquares <- sciMelt[sciMelt$value > 1.128380, ]
head(sciNoSquares)
head(sciMelt)
tail(sciMelt)
glimpse(sciMelt)
sciTRASH <- melt(sci_DF, id = c("treatment", "year"), measure.vars = "sci")
head(sciTRASH)
# Plot sci values for each treatment, grouped by plot, as a function of year
sciSmoothed_plot <- ggplot(sciNoSquares[sciNoSquares$treatment == "unun", ], aes(x = as.numeric(year), y = value, color = plot, group = plot)) +
    geom_smooth() +
    ggtitle("Shape Complexity Index Timeseries") +
    labs(x = "Year Index", y = "Shape Complexity Index") #+
    #facet_wrap(~treatment)
sciSmoothed_plot

# Density plots of sci by treatment group
sciDensity_title <- c("Shape complexity density by plot and grazing treatment")
sciDensity_plot <- ggplot(sciNoSquares[sciNoSquares$treatment == "unun", ], aes(value, colour = year, group = year)) +
    geom_density() +
    ggtitle(sciDensity_title) +
    labs(x = "Shape Complexity Index Value", y = "Density") +
    facet_wrap(~plot)
sciDensity_plot

# ln Transformed Density plots of sci by treatment group
sciTransfDens_plot <- ggplot(sciNoSquares, aes(value, colour = plot, group = plot)) +
    geom_density() +
    scale_x_continuous(trans = "logit") +
    ggtitle("Shape Complexity Index by Plot and Grazing Treatment") +
    labs(x = "Transformed Shape Complexity Index", y = "Density (%)") +
    facet_wrap(~treatment)
sciTransfDens_plot

sciDensityByYear_plot <- ggplot(sciMelt) +
    geom_density(aes(x = value, y = ..density.., colour = treatment, group = year)) +
    ggtitle("Shape Complexity Density by Year and Treatment") +
    labs(x = "Shape Complexity Index Value", y = "Density (%)") +
    facet_wrap(~treatment)
sciDensityByYear_plot

# Violin plots
sciViolinByTreatment_plot <- ggplot(sciMelt) +
    geom_violin(aes(x = year, y = value, colour = treatment, group = year)) +
    ggtitle("Shape Complexity by Year") +
    labs(x = "Year", y = "Shape Complexity Index") +
    facet_wrap(~treatment)
sciViolinByTreatment_plot

# Scatter plots
sciScatterByTreatment_plot <- ggplot(sciMelt) +
    geom_point(aes(x = year, y = value, colour = treatment, group = treatment)) +
    ggtitle("Shape Complexity Inex by Year and Grazing Treatment") +
    labs(x = "Year Index", y = "Shape Complexity Index") +
    facet_wrap(~treatment)
sciScatterByTreatment_plot
```   

