---
title: "Shortgrass Steppe Point Pattern Analysis"
author: "Darin Schulte"
date: "October 2, 2016"
output:
  html_document: 
    toc: true
    toc_depth: 6
---

### Description   
Data processing associated with investigating the application of point pattern analysis techniques on polygon data. Work is divided into three main components:   

**Component**   |       **Purpose**     
------------|---------------
PPA         |   Traditional Point Pattern Analysis
Polygon PPA |   Polygon based pair-correlation analysis    
Aerial Polygon PPA   |   Polygon based analysis accounting for area of neighbors   

### General terms and descriptions   
**Concept**   |       **Description**     
------------|---------------
Pair Correlation Function $g(r)$       |   Based on object to object distances and describes the aggregation and regularity at a given radius $r$. A completely random pattern: $g(r) = 1$. Clustering: $g(r) > 1$. Regularity: $g(r) < 1$. $g(r)$ can take any value between $0$ and $infty$ and generally approaches $1$ as $r$ increases.
Polygon PPA |   Polygon based pair-correlation analysis    
Aerial Polygon PPA   |   Polygon based analysis accounting for area of neighbors      


###Load Necessary Packages      
```{r}
# Spatial data and analysis packages
library(rgdal)
getGDALVersionInfo()
library(gdalUtils)
library(tools)
library(raster)
library(sp)
library(shapefiles)
library(reshape2)
library(spatstat)
library(maptools)
library(rgeos)

# Plotting packages
library(ggplot2)
library(grid)
library(Rmisc)
library(GGally)

# General packages for data manipulation and statistics
library(data.table)
library(dplyr)
library(stats)
library(plyr)
library(cleangeo)
#library(car)
#library(lme4)
#library(nlme)
```   

<br/>   

-----------    

### Set working directory    
Files were created in the *idaho_SIanalysis.RMD* file.   
```{r}
setwd("C:\\Users\\darin\\Dropbox\\wd_R\\idaho_analysis\\working_data")
getwd()

```   


### Read in shapefile     
Current version of the code will not perform randomization correctly if a polygon intersects more than one bounding box edge.    
```{r}
# Read in the files
polyFile <- readShapePoly("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/cleanedShapeFiles/poly_unun_7_1998_cleaned.shp")
plot(polyFile)

# Remove duplicate SP_ID column from @data. NOTE: @data ID does not track with the SPDF Polygon ID's used below 
polyFile@data <- polyFile@data[, c(2, 4)]

```    

### Convert polygons to points and create a ppp object    
```{r}
# Get centroids for the polygons
#NOTE 1 pt is outside its polygon using this method. Here, points are not forced to be within polygons.
centPoints <- gCentroid(polyFile, byid=TRUE)
plot(centPoints, col="skyblue", add = TRUE)  

# Create ppp object from centPoints
centPoints <- as.ppp(centPoints)
plot(centPoints)
```   

### PPA - Density plot     
```{r}
# Plot smoothed density
plot(density(centPoints), main = "Density Plot")
```   

### PPA - Nearest Neighbor Distances         
```{r}
# Calculate distance from points to their nearest neighbors
nnEst <- Gest(centPoints)
nnEst
plot(nnEst) # Values below G_pois suggest regularity

# Get simulation envelopes for Gest (nrank sets the rank of the hi and low values)
nnEst_env <- envelope(centPoints, Gest, nsim = 500, nrank = 13)
plot(nnEst_env) # Values below G_pois suggest regularity

#_________________________________
# Mean Nearest Neighbor Distance
meanNND <- nndist(centPoints)
hist(meanNND)
summary(meanNND)
# Quick comparison to CSF of same intensity
mnnd <- mean(meanNND)
exp_nnd <- 0.5 / sqrt(centPoints$n/area(centPoints))
mnnd/exp_nnd
```      

### PPA - Ripley's K & L-function            
```{r}
# Ripley's K-function
ripK_Est <- Kest(centPoints)
ripK_Est
plot(ripK_Est) # Values below K_pois suggest regularity

# Get simulation envelopes for Kest
ripK_Est_env <- envelope(centPoints, Kest, nsim = 500, nrank = 13)
plot(ripK_Est_env) # Values below K_theo suggest regularity

# L-function: Transformation of K-function helpful for visualization
l_Est <- Lest(centPoints)
l_Est
plot(l_Est) # Values below L_pois suggest regularity (same interpretation as K-function)

# Get simulation envelopes for Lest (nrank sets the rank of the hi and low values)
l_Est_env <- envelope(centPoints, Lest, nsim = 500, nrank = 13)
plot(l_Est_env) # Values below L_theo suggest regularity (same interpretation as K-function)
```      

### PPA - Pair-Correlation function g(r)  
Pair correlation function tells you whether there are more (above envelope) or fewer (below envelope) pairs of events at the given range of spacings than expected to occur by chance.   
```{r}
# Pair-correlation function
pcf_Est <- pcf(centPoints)
plot(pcf_Est)

# Get simulation envelopes for PCF (nrank sets the rank of the hi and low values)
pcf_Est_env <- envelope(centPoints, pcf, nsim = 500, nrank = 13)
plot(pcf_Est_env) # Values below L_theo suggest regularity (same interpretation as K-function)
# Change the range of y axis for better visualization
plot(pcf_Est_env, ylim=c(0, 5))
```      


