---
title: "Shortgrass Steppe Polygon Pattern Analysis"
author: "Darin Schulte"
date: "October 2, 2016"
output:
  html_document: 
    toc: true
    toc_depth: 6
---

### Description   
Data processing associated with investigating the application of point pattern analysis techniques on polygon data. Work is divided into three main components:   

**Component**   |       **Purpose**     
------------|---------------
PPA         |   Traditional Point Pattern Analysis
Polygon PPA |   Polygon based pair-correlation analysis    
Aerial Polygon PPA   |   Polygon based analysis accounting for area of neighbors   

### General terms and descriptions   
**Concept**   |       **Description**     
------------|---------------
Pair Correlation Function $g(r)$       |   Based on object to object distances and describes the aggregation and regularity at a given radius $r$. A completely random pattern: $g(r) = 1$. Clustering: $g(r) > 1$. Regularity: $g(r) < 1$. $g(r)$ can take any value between $0$ and $infty$ and generally approaches $1$ as $r$ increases.
Polygon PPA |   Polygon based pair-correlation analysis    
Aerial Polygon PPA   |   Polygon based analysis accounting for area of neighbors      


###Load Necessary Packages      
```{r}
# Spatial data and analysis packages
library(rgdal)
getGDALVersionInfo()
library(gdalUtils)
library(tools)
library(raster)
library(sp)
library(shapefiles)
library(reshape2)
library(spatstat)
library(maptools)
library(rgeos)

# Plotting packages
library(ggplot2)
library(grid)
library(Rmisc)
library(GGally)

# General packages for data manipulation and statistics
library(data.table)
library(dplyr)
library(stats)
library(plyr)
library(cleangeo)
#library(car)
#library(lme4)
#library(nlme)
```   

<br/>   

-----------    

### Set working directory    
Files were created in the *idaho_SIanalysis.RMD* file.   
```{r}
setwd("C:\\Users\\darin\\Dropbox\\wd_R\\idaho_analysis\\working_data")
getwd()
```   


### Read in shapefile     
Current version of the code will not perform randomization correctly if a polygon intersects more than one bounding box edge.    
```{r}
# Read in the files
polyFile <- readShapePoly("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/cleanedShapeFiles/poly_unun_7_1998_cleaned.shp")
plot(polyFile)
# Remove duplicate SP_ID column from @data. NOTE: @data ID does not track with the SPDF Polygon ID's used below 
polyFile@data <- polyFile@data[, c(2, 4)]
polyFile@polygons[[1]]@Polygons
```  


## Polygon-based pattern analysis     

### Identify polygons that intersect the edge of the plot   
The following is working with code from Ben Marwick as a base. See _https://github.com/benmarwick/Rocks-and-burials-at-Madjebebe/blob/master/code/Rocks-and-burials-at-MJB.Rmd_ for more details.      

Create a polygon of the bounding box for the entire plot   
```{r}
# Create a polygon of the bounding box of the entire plot
bboxPolygon <- as(extent(polyFile), 'SpatialPolygons')
plot(bboxPolygon, add= TRUE)

# Special case: some polygons clearly on the edge do not quite touch the bbox. Try creating a 1 cm inner buffer for the bounding box to bring it in a bit.
bboxPolygon <- gBuffer(bboxPolygon, width = -0.01, capStyle = "SQUARE")
bboxPolygon
```      

Convert the bounding box to a `SpatialLines` object   
```{r}
# Convert bboxPolys to a SpatialLines object 
bboxAsLines <- as(bboxPolygon, "SpatialLines")
bbLinesList <- unlist(bboxAsLines@lines[[1]])
lineNames <- c("line1", "line2", "line3", "line4")
# Create SpatialLines objects for the bbox lines
bbLines_1 <- SpatialLines(list(Lines(list(Line(bbLinesList@Lines[[1]]@coords[1:2, ])), ID = lineNames[[1]])))
bbLines_2 <- SpatialLines(list(Lines(list(Line(bbLinesList@Lines[[1]]@coords[2:3, ])), ID = lineNames[[2]])))
bbLines_3 <- SpatialLines(list(Lines(list(Line(bbLinesList@Lines[[1]]@coords[3:4, ])), ID = lineNames[[3]])))
bbLines_4 <- SpatialLines(list(Lines(list(Line(bbLinesList@Lines[[1]]@coords[4:5, ])), ID = lineNames[[4]])))
# Join these into a single SpatialLines object
bbLines <- bbLines_1
bbLines <- spRbind(bbLines, bbLines_2)
bbLines <- spRbind(bbLines, bbLines_3)
bbLines <- spRbind(bbLines, bbLines_4)
# Clean up old objects
rm(bbLines_1, bbLines_2, bbLines_3, bbLines_4, lineNames)
```   

Create a list that stores the index of the `bbLines` object where an overlap with a given polygon occurs   
```{r}
# Get list of the bounding box edge index of overlap
edgePolys <- sp::over(polyFile, bbLines, byid = TRUE)
```   


### Generate random points within the bounding box.     
**Note** the number of points returned is approximately the value given. Better to generate extra points.    
```{r}
# Get the number of polygons in the shapefile
numPoints <- length(polyFile) 
rndPoints <- spsample(bboxPolygon, numPoints*2, type = "random")
#rndPoints <- spsample(bboxPolygon, 200, type = "random")
plot(rndPoints)
#plot(bboxPolygon, add = TRUE, border = "lightsteelblue4")

# Define a function to remove points if they are overlapped by a newly placed polygon
removePoints <- function(overPoly, points2check) {
    overPoints <- sp::over(overPoly, points2check, returnList = TRUE)
    if (length(overPoints[[1]] != 0)) {
        newPoints <- points2check[-overPoints[[1]]]
    } else {
        newPoints <- points2check
    }
    return(newPoints) 
}
# overPointsTRASH <- sp::over(polyFile[poly_index[39], ], rndPoints, returnList = TRUE)
# length(overPointsTRASH[[1]]) != 0
```        

### Randomly rotating and moving the polygons   

Defines the process by which each plant polygon is randomly rotated and shifted to a point from the `rndPoints` object, or, if it touches the bounding box of the plot, it is rotated by a multiple of 90 degrees and assigned to a random location along one of the other edges. This creates a shapefile of the randomly shuffled polygons within the bounding box. This is repeated to generate a user defined number of shapefiles.     

Code originally based on _https://github.com/benmarwick/Rocks-and-burials-at-Madjebebe/blob/master/code/Rocks-and-burials-at-MJB.Rmd_, however, recent updates to several packages (most notably the raster package) made the process much more streamlined.   
```{r}
# Create list of polygons sorted by size including whether or not they intersect the bounding box (they will be placed largest to smallest) 
polys_sorted <- data.frame(id = integer(), intersectEdge = integer(), area = double()) 
for (areas in 1:length(polyFile)) {
    polys_sorted[areas, 1] <- polyFile@polygons[[areas]]@ID
    polys_sorted[areas, 2] <- edgePolys[[areas]]
    # May be better to use gArea, but since the exact area is not really the focus, it may not be necessary.
    polys_sorted[areas, 3] <- polyFile@polygons[[areas]]@area
}
# Sort by descending area
polys_sorted <- polys_sorted[order(-polys_sorted[,3]),]

# The row names correspond to the polygon index (ID's start at 0 so row numbers are ID + 1)
poly_index <- row.names(polys_sorted)
poly_index <- as.numeric(poly_index)

# Create list to store the output from each full run as a list of SpatialPolygonsDataFrames
# The value supplied to 'length' will determine the number of random patterns generated
#   Perry et al 2006 pg 62 note Diggle 2003 in saying 500 replicats seems appropriate for (alpha = 0.01) in
#   tests of statistical significance
simsList <- vector("list", length = 500)   # Change back to 500

# Create a list of random angles (multiple of 90 degrees) by which to rotate the polygon (around its centroid).
# Here, a list of length(number of polygons in the dataset) is created for the hypothetical case in which all polygons
#   are touching an edge. This is never the case, but it makes sure that enough numbers are generated so that indexing 
#   by 'plant' can be used to get the random rotation angle.
# NOTE: rotates clockwise if 'by' is positive
#edgePolyRotAngle <- sample(seq(0, 270, by = 90), size = length(polyFile), replace = TRUE)


# The same process but for non-edge polygons. This creates a list of random angles from 0 to 359 of length(number of 
#   polygons in the dataset) in the event that no polygons are touching an edge
#rotateAngle <- sample(seq(0, 359, by = 1), size = length(polyFile), replace = TRUE)

# Define an empty vector of length 2 to store the centroid coordinates for a given polygon
centCoords <- vector(mode = "double", length = 2)

# Function to calculate the centroid coordinates for a given polygon
getCentCoords <- function(plantMoving) {
    plantCentCoords <- slot(slot(slot(plantMoving, "polygons")[[1]], "Polygons")[[1]], "labpt")
    return(plantCentCoords)
}

# Define empty vector of length 2 to store the random point coordinates which the plant will move to 
rn <- vector(mode = "double", length = 2)

# Define a list of one element (a vector of length 2) for storing offset values
offsetValue <- list(vector(mode = "double", length = 2))

# Define a function for calculating the offset values
getOffsetVal <- function(coordVect, randomPoint) {
    o_s_val <- coordVect - randomPoint
    o_s_val <- list(-o_s_val)
    return(o_s_val)
}

# Create an empty list to store results of checking for overlap
# NOTE: Not run in the case where overlap is allowed
overlapCheck = list()

# Create an empty vector of length 1 to hold boolean value for crosses with other polygons
noCrosses <- vector(mode = "logical", length = 1)

# Main Simulation Code
for (i in 1:length(simsList)) {
    # Track number of attempts to place a polygon (breaks at 1000 tries)
    repTracker <- 1
    repMax <- 2000
    # Variable for query index
    plant <- 1
    plantMax <- length(poly_index) + 1

    repeat {
        # Get id of plants from the sorted list
        plantToMoveID <- poly_index[plant]  

        # Select plant to rotate and move
        plantToMove <- polyFile[plantToMoveID, ]
        #plot(plantToMove)
        #polyFile[114, ]

        # If the polygon is touching the bounding box,
        #   1) rotate it by a multiple of 90 degrees, 
        #   2) find the edge it intersects with, and 
        #   3) move to a random position on the (different) appropriate bounding box edge
        if (!is.na(polys_sorted[plant, 2])) {
            # get the coordinates of the labpt, labpt, bbox also (centroid coords and bounding box)
            centCoords <- getCentCoords(plantToMove)

            # Get a random angle (multiple of 90 degrees) by which to rotate the polygon (around its center)
            # NOTE: rotates clockwise if 'by' is positive
            edgePolyRotAngle <- sample(seq(0, 270, by = 90), size = 1)   
            
            # Rotate the polygon
            plantToMove <- elide(plantToMove, rotate = edgePolyRotAngle, center = centCoords)
            
            # Get the coordinates for plantToMove bbox (after rotation)
            plantToMoveBBox <- plantToMove@bbox
            
            # Get the dimensions of the bbox (after rotation)
            plantToMoveBBoxWidth <- abs(plantToMoveBBox[1,1] - plantToMoveBBox[1,2])  # in the x dimension
            plantToMoveBBoxHeight <- abs(plantToMoveBBox[2,1] - plantToMoveBBox[2,2]) # in the y dimension
            
            # Get the shortest distance between the plantToMoveBBox and all edges
            moveDistance <- gDistance(plantToMove, bbLines, byid = TRUE)

            # Generate random value to assign to a point coordinate along an edge depending on the size of the polygon
            randXOffset <- runif(1, (0.5 * plantToMoveBBoxWidth), (1 - 0.5 * plantToMoveBBoxWidth))
            randYOffset <- runif(1, (0.5 * plantToMoveBBoxHeight), (1 - 0.5 * plantToMoveBBoxHeight))
       
            # Move the polygon to the appropriate edge depending on its rotation
            if (polys_sorted[plant, 2] == 1) { # If polygon is initially touching the first side
                if (edgePolyRotAngle == 0) {
                    offsetCoord_x <- 0.0
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                    #gContains(bboxPolygon, plantMoved) == TRUE
                } else if (edgePolyRotAngle == 90) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- moveDistance[2]
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 180) {
                    offsetCoord_x <- moveDistance[3]
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 270) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- -(moveDistance[4])
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                }
            } else if (polys_sorted[plant, 2] == 2) { # If polygon is initially touching the 2nd side
                if (edgePolyRotAngle == 0) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- 0.0
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 90) {
                    offsetCoord_x <- moveDistance[3]
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 180) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- -(moveDistance[4])
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 270) {
                    offsetCoord_x <- -(moveDistance[1])
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                }
            } else if (polys_sorted[plant, 2] == 3) { # If polygon is initially touching the 3rd side
                if (edgePolyRotAngle == 0) {
                    offsetCoord_x <- 0.0
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 90) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- -(moveDistance[4])
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 180) {
                    offsetCoord_x <- -(moveDistance[1])
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 270) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- moveDistance[2]
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                }
            } else if (polys_sorted[plant, 2] == 4) { # If polygon is initially touching the 4th side
                if (edgePolyRotAngle == 0) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- 0.0
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 90) {
                    offsetCoord_x <- -(moveDistance[1])
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 180) {
                    offsetCoord_x <- -(centCoords[1] - randXOffset)
                    offsetCoord_y <- moveDistance[2]
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                } else if (edgePolyRotAngle == 270) {
                    offsetCoord_x <- moveDistance[3]
                    offsetCoord_y <- -(centCoords[2] - randYOffset)
                    plantMoved <- elide(plantToMove, shift = c(offsetCoord_x, offsetCoord_y))
                }
            }
        } else {
            # get the coordinates of the labpt, labpt, bbox also (centroid coords and bounding box)
            centCoords <- getCentCoords(plantToMove)
            
            # rotate the plant by a random angle
            rotateAngle <- sample(seq(0, 359, by = 1), size = 1)
            plantToMove <- elide(plantToMove, rotate = rotateAngle, center = centCoords) # NOTE: rotates clockwise if 'by' is positive
            
            # get a random point in the overall bounding box to move to
            samplePoint <- sample(rndPoints, size = 1, replace = TRUE)  #### Was replace = FALSE
            rn <- unname(samplePoint@coords)
            #rn <- unname(rndPoints@coords[sample(1:length(rndPoints), 1),])
            
            # shift all vertices coordinates
            offsetValue <- getOffsetVal(centCoords, rn)

            # shift the spatialPolygons object
            plantMoved <- elide(plantToMove, shift = c(offsetValue[[1]][1], offsetValue[[1]][2]))
        }
        
        # Test to see whether the shifted polygon is within the plot bounding box (allowing for intersection)
        if (gContains(bboxPolygon, plantMoved) == TRUE) {
            if (plant == 1) {
                plants_rnd <- plantMoved
                plant <- plant + 1
                # Remove points overlapped by the placed polygon
                rndPoints <- removePoints(plantMoved, rndPoints)
                #rndPointsOver <- sp::over(plantMoved, rndPoints, returnList = TRUE)
                #rndPoints <- rndPoints[-rndPointsOver[[1]]]
                #cat('\n', 'Moving on to plant ', plant)
                plot(polyFile[poly_index[39], ], add = TRUE)
            } else {
                # Check to see whether it overlaps with any other polygons
                overlapCheck <- sp::over(plantMoved, plants_rnd, returnList = TRUE)
                noCrosses <- (nrow(overlapCheck[[1]]) == 0) || is.null(nrow(overlapCheck[[1]]))

                if (noCrosses == TRUE) {
                    # Merge this SPDF object with plants_rnd
                    plants_rnd <- rbind(plants_rnd, plantMoved)
                    plant <- plant + 1
                    # Remove points overlapped by the placed polygon
                    rndPoints <- removePoints(plantMoved, rndPoints)                  
                    #cat('\n', 'Moving on to plant ', plant, ' rndPoints = ', length(rndPoints))
                    #plot(plants_rnd)
                }
            }
        } else {
            plant <- plant
            #cat('\n', "Failed try. Staying at plant ", plant)
            repTracker <- repTracker + 1
            #cat('\n', 'Attempt number ', repTracker)
        }
        
        # Check for number of tries
        if (repTracker == repMax) {
            cat('\n', 'Max reps reached. Exiting this attempt.')
            break
        }
        if (plant == plantMax) {
            #cat('\n', 'That should be all of the plants')
            cat('\n', "#######  COMPLETING ROUND ###############", i)
            simsList[[i]] <- plants_rnd
            rndPoints <- spsample(bboxPolygon, numPoints*2, type = "random")
            break
        }
    }  # End of repeat
    
}  # End of for loop

plot(simsList[[1]])
plot(simsList[[300]])
plot(polyFile)
plot(bboxPolygon, add = TRUE)
polys_sorted

# Verify that each has the correct number of polygons
testLength <- lapply(simsList, function(x) {length(x)})
lapply(simsList, function(x) {length(x) == testLength})

#___________________________________________________________________________
# Write to file
# Save simsList as an RData file for loading in the future
saveRDS(simsList, "C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/unun_7_1998_simsList.RDS")
```     


## Buffering and overlap calculations   
```{r}
# Read in the simsList list of SPDFs (created in the previous code chunk)
simsList <- readRDS("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/unun_7_1998_simsList.RDS")

# Read in the original shapefile (for which simsList patterns were generated)
plotTest <- readOGR("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/cleanedShapeFiles/poly_unun_7_1998_cleaned.shp", layer = "poly_unun_7_1998_cleaned")
plot(plotTest)

# Append the SPDF for the original shapefile to simsList as the last element
simsList[[501]] <- plotTest

#___________________________________________________
# Buffer testing
# Buffer testing requires shapefiles to be read in with readOGR to create inner buffers. Write simsList to shapefiles and read in again with readOGR.
#dir.create("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/simsLists")
#dir.create("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/simsLists/unun_7_1998_simsList")

# Create list of file names of 'cleaned' polygons(just the .shp file names)
clnShpFiles <- list.files("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/cleanedShapeFiles", pattern = glob2rx("*.shp"), full.names = TRUE)

# Get only those names of plots that were simulated in simsList
simFile <- clnShpFiles[[260]]
simBaseName <- basename(simFile)
simBaseName <- file_path_sans_ext(simBaseName)
simIndex <- seq(from = 1, to = length(simsList), by = 1)
simNames <- lapply(simIndex, function(x) paste("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/simsLists/unun_7_1998_simsList/", simBaseName, "_sim_", simIndex[x], sep=""))

# Write simulations to shapefiles
lapply(1:length(simsList), function(x) {
    shpToWrite <- simsList[[x]]
    shpToWriteName <- simNames[[x]]
    writeSpatialShape(shpToWrite, shpToWriteName)
})
# Clean up environment variables
#rm(clnShpFiles, simFile, simIndex, simNames, simsList)


#________________________________________________________
# Create list of file names of 'simsList' polygons(just the .shp file names)
simShpFiles <- list.files("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/simsLists/unun_7_1998_simsList", pattern = glob2rx("*.shp"), full.names = TRUE)  

# The list is sorted alphabetically, so simulation numbers are not in correct order. Correct for that.
# Source for this process: http://stackoverflow.com/questions/10777367/how-can-i-read-the-files-in-a-directory-in-sorted-order-using-r
# Split between the part that comes before the numeric index and the "index.shp"
split <- strsplit(simShpFiles, "C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/simsLists/unun_7_1998_simsList/poly_unun_7_1998_cleaned_sim_")
# Strip the "1.shp" etc. to get just the numeric values and turn the characters to numeric type
split <- as.numeric(sapply(split, function(x) x <- sub(".shp", "", x[2])))

# Sort by numeric order
simShpFiles <- simShpFiles[order(split)]
simLayer <- basename(simShpFiles)
simLayer <- file_path_sans_ext(simLayer)

# Create list to store the distances of the buffers
bufferDists <- seq(from = 0.0, to = 0.20, by = 0.01)

# Buffering
bufferTesting <- function(simFile) {
    # Get the layer name for the simFile
    simLayerName <- basename(simFile)
    simLayerName <- file_path_sans_ext(simLayerName)
    
    # Read in the shapefile
    plotTest <- readOGR(simFile, layer = simLayerName)
#plotTest <- readOGR(simShpFiles[[52]], layer = simLayer[[52]])
#plot(plotTest)
    # Get the bounding box of the plot
    bboxPolygon <- as(extent(plotTest), "SpatialPolygons")
    
    # Inset bbox for edge effect 
    ################TODO: TEST FOR MULTIPLE INNER BUFFER SIZES###############
    inBBoxPolygon <- gBuffer(bboxPolygon, width = -0.1, capStyle = "SQUARE")

    # To deal with polygons that touch the bounding box edge but also extend into the inner bbox buffer
    #________________________________________________________
    # 1) get IDs of polygons not intersecting the plot bbox
    notEdgeIDs <- sapply(lapply(slot(plotTest[c(gContainsProperly(bboxPolygon, plotTest, byid = TRUE)), ], "polygons"),
                             function(x) SpatialPolygons(list(x))), function(x) slot(slot(x, "polygons")[[1]], "ID"))
#plot(plotTest[notEdgeIDs, ], col = "skyblue")    
    #________________________________________________________
    # 2) Get IDs from notEdgeIDs polygons that are contained within the inner buffer of the bounding box allowing for intersection 
    innerPolys <- plotTest[notEdgeIDs, ]
    innerPolys <- sapply(lapply(slot(innerPolys[c(gIntersects(inBBoxPolygon, innerPolys, byid = TRUE)), ], "polygons"),
                            function(x) SpatialPolygons(list(x))), function(x) slot(slot(x, "polygons")[[1]], "ID"))
#plot(bboxPolygon)
#plot(inBBoxPolygon, add = TRUE)
#plot(plotTest[innerPolys, ], col = "skyblue", add = TRUE)
    #_________________________________________________________
    # 3) NOTE: Buffering performed on plotTest[innerPolys, ]

    # Get the @data for plotTest
    # polyData <- plotTest@data
    polyData <- plotTest[innerPolys, ]@data

    # Buffer each polygon (returns list of SPDFs one for each buffer distance)
    # Note: first buffer is set to 0.00 so it is the polygons without a buffer
    # buffTest <- lapply(1:length(bufferDists), function(x) {
    #     gBuffer(plotTest, width = bufferDists[x], byid = TRUE)
    # })
    buffTest <- lapply(1:length(bufferDists), function(x) {
        gBuffer(plotTest[innerPolys, ], width = bufferDists[x], byid = TRUE)
    })
#plot(buffTest[[1]])
    # Get the buffer rings (i.e., just the buffer ring and not the original polygon (or previous rings))
    diffTest <- lapply(2:length(buffTest), function(x) {
        lapply(1:length(buffTest[[x]]), function(row) gDifference(buffTest[[x]][row, ], buffTest[[(x-1)]][row, ]))
    })    
#plot(diffTest[[1]][[1]], col = "skyblue")
    # All of the IDs for individual rings are set to "1". Create list of ID values and assign them to the diffTest polys
    polyIDs <- as.character(seq(from= 0, to = length(plotTest)-1, by = 1))
    
    # Merge the diffTest (list of, lists of SpatialPolygons) to get a list of SPDF objects
    joined <- lapply(1:length(diffTest), function(x) {
        SpatialPolygons(lapply(1:length(diffTest[[x]]), function(i) {
            Pol <- slot(diffTest[[x]][[i]], "polygons")[[1]]
            slot(Pol, "ID") <- polyIDs[i]
            Pol
        }))
    })
    joinedSPDF <- lapply(1:length(joined), function(x) {
        SpatialPolygonsDataFrame(Sr = joined[[x]], data = polyData, FALSE)
    })
#plot(joinedSPDF[[1]])
#plot(joinedSPDF[[2]], col = "green", add = TRUE)
    # Clip buffers to the bounding box of the plot
    buffClipped <- lapply(1:length(joinedSPDF), function(x) {
        gIntersection(bboxPolygon, joinedSPDF[[x]], byid = TRUE, drop_lower_td = TRUE)
    })
#plot(buffClipped[[20]])
    ## Left as reference for creating list of polygons and accessing ID slot
    # listOfPolys_2 <- lapply(1:length(buffClipped_2), function(x) {
    #     lapply(slot(buffClipped_2[[x]], "polygons"), function(x) SpatialPolygons(list(x)))
    # })    
    # polyIDs_2 <- lapply(1:length(listOfPolys_2), function(x) {
    #     sapply(listOfPolys_2[[x]], function(x) slot(slot(x, "polygons")[[1]], "ID"))
    # })
    
    # Get the area of the ring buffers
    ringArea <- lapply(1:length(joinedSPDF), function(x) {
        gArea(joinedSPDF[[x]], byid = TRUE)
    })
    ringAreaDF <- as.data.frame(ringArea)
    #head(ringAreaDF)
    colnames(ringAreaDF) <- paste0("b_", c(1:20), "cm")
#ringAreaDF[, 1]
    # Get proportion of diffTest buffer rings that is / are within the bounding box
    propRingIn <- lapply(1:length(joinedSPDF), function(x) {
        sapply(1:length(joinedSPDF[[x]]), function(row) gArea(buffClipped[[x]][row, ])/gArea(joinedSPDF[[x]][row, ]))
    })
    propRingInDF <- as.data.frame(propRingIn)
#propRingInDF[, 1]
    # Assign row and column names to buffDF (Row names code not run - 10.03.16)
    colnames(propRingInDF) <- paste0("b_", c(1:20), "cm")
    # rownames(propRingInDF) <- row.names(joinedSPDF[[1]])
    
    # Get the ID's of the plotTest polygons that the given buffer overlaps. (Note, using gIntersection returns
    # an ID for self intersection. Using gOverlaps returns NULL for self intersection only, and plotTest ID when
    # the buffer overlaps with a plotTest polygon)
    overlapsTest <- lapply(1:length(joinedSPDF), function(x) {
        gOverlaps(joinedSPDF[[x]], plotTest, byid = TRUE, returnDense = FALSE)
    })
    
    # # Old version (order of plotTest & joinedSPDF is incorrect)
    # overlapsTest <- lapply(1:length(joinedSPDF), function(x) {
    #      gOverlaps(plotTest, joinedSPDF[[x]], byid = TRUE, returnDense=FALSE)  
    # })

# overlapsTest_2[[1]][[2]]
# plot(joinedSPDF[[1]][2, ])
# joinedSPDF[[1]][2, ]
# plot(plotTest, add = TRUE)
# plotTest[71, ]
# plotTest[overlapsTest_2[[1]][[2]][[3]], ]
# plot(plotTest[overlapsTest_2[[1]][[2]], ])
    
    # Get the intersection polygons between buffer rings and plotTest polygons 
    intersectTest <- lapply(1:length(joinedSPDF), function(x) {
        lapply(1:length(joinedSPDF[[x]]), function(j) {
            if (!is.null(overlapsTest[[x]][[j]])) {
                gIntersection(joinedSPDF[[x]][j, ], plotTest[overlapsTest[[x]][[j]], ], byid = TRUE)
            }
        })
    })
    
# # Testing version (left as reference)
# intersectTest_2 <- lapply(1:length(joinedSPDF), function(x) {
#     lapply(1:length(joinedSPDF[[x]]), function(j) {
#         if (!is.null(overlapsTest_2[[x]][[j]])) {
#             gIntersection(joinedSPDF[[x]][j, ], plotTest[overlapsTest_2[[x]][[j]], ], byid = TRUE)
#         }
#     })
# })

    # Get the area of the intersections
    intersectAreas <- lapply(1:length(intersectTest), function(x) {
        lapply(1:length(intersectTest[[x]]), function(j) {
            if (!is.null(intersectTest[[x]][[j]])) {
                gArea(intersectTest[[x]][[j]])
            }
        })
    })
    
# Testing version (left as reference)    
# intersectAreas_2 <- lapply(1:length(intersectTest_2), function(x) {
#     lapply(1:length(intersectTest_2[[x]]), function(j) {
#         if (!is.null(intersectTest_2[[x]][[j]])) {
#             gArea(intersectTest_2[[x]][[j]])
#         }
#     })
# })

    # Convert intersectAreas to numeric data type(s) to deal with NULL values (Warnings are about the NAs and can be ignored)
    intDF <- data.frame(sapply(intersectAreas, function(x) as.numeric(as.character(x))))
# # Testing version (left as reference)    
# intDF_2 <- data.frame(sapply(intersectAreas_2, function(x) as.numeric(as.character(x))))    
    
    # Calculate the area version of intensity (lambda) for use in calculating wgtTest below
    plotIntensity <- gArea(plotTest)/gArea(bboxPolygon)

    # Calculations for area weighted g(r)
    wgtTest <- intDF / (plotIntensity * propRingInDF * ringAreaDF)
# # Testing version (left as reference)     
# wgtTest_2 <- intDF_2 / (plotIntensity * propRingInDF * ringAreaDF)    

    # Means for each buffer ring distance
    obsMeans <- (colSums(wgtTest, na.rm = TRUE))/nrow(wgtTest)
# # Testing version (left as reference)     
# obsMeans_2 <- (colSums(wgtTest_2, na.rm = TRUE))/nrow(wgtTest_2)    
    
    return(obsMeans)
}

# Run the bufferTesting function (duration ~ 3 hours for unun_7_1998)
buffDF <- do.call("rbind", lapply(simShpFiles, function(j) bufferTesting(j)))

# Write to .rds and .csv files
saveRDS(buffDF, file="C:\\Users\\darin\\Dropbox\\wd_R\\idaho_analysis\\working_data\\msData\\Q20_56_C_bufferDF.rds")
saveRDS(buffDF, file="C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_bufferDF_2.rds")
write.csv(buffDF, file="C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_bufferDF_2.csv")
```   

```{r}
# Read in the results of the bufferTesting function (note: it is read in as a matrix)
bufferDF <- readRDS("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_bufferDF_2.rds")

# Correction for polygon approach (Nuske et al. 2009)
simResults <- bufferDF[1:500, ]
simMeans <- colMeans(simResults)
simResults_Norm <- simResults %*% diag(1/simMeans)

# Apply normailization to field data NOTE: NEED TO VERIFY THIS IS THE CORRECT WAY TO DO THIS
field_Norm <- bufferDF[501, ] %*% diag(1/simMeans)

# Transpose and melt to long format for ggplot2
simNorm_T <- t(simResults_Norm)
head(simNorm_T)
field_Norm_T <- t(field_Norm)

# Coerce simNorm_T & field_Norm_T matrices to dataframes
simNormDF_T <- as.data.frame(simNorm_T)
bufferDists <- seq(from = 0.0, to = 0.20, by = 0.01)
simNormDF_T <- cbind(simNormDF_T, bufferDists[2:21])
colnames(simNormDF_T) <- c(paste0("sim_", seq(from = 1, to = 500, by = 1), sep = ""), "buffer")
fieldNormDF_T <- as.data.frame(field_Norm_T)
fieldNormDF_T <- cbind(fieldNormDF_T, bufferDists[2:21])
colnames(fieldNormDF_T) <- c("g_r_field", "buffer_field")

# Convert to data.table
setDT(simNormDF_T)
meltedsimNorm <- melt(simNormDF_T, id.vars="buffer")
setDT(fieldNormDF_T)

# Write to .rds 
saveRDS(meltedsimNorm, "C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_meltedSimNorm_2.rds")
write.csv(meltedsimNorm, "C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_meltedSimNorm_2.csv")
saveRDS(fieldNormDF_T, "C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_fieldNormDT_T_2.rds")

# Read in .rds 
meltedsimNorm <- readRDS("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_meltedSimNorm_2.rds")
head(meltedsimNorm)
tail(meltedsimNorm)
is.data.table(meltedsimNorm)
fieldNormDF_T <- readRDS("C:/Users/darin/Dropbox/wd_R/sgs_spatial/sgs_data/sgs_unun_7_1998_fieldNormDT_T_2.rds")
head(fieldNormDF_T)
is.data.table(fieldNormDF_T)

# Change the types of column data
glimpse(meltedsimNorm)
names(meltedsimNorm) <- c("b_dist", "sim", "g_r")
meltedsimNorm$g_r <- unlist(meltedsimNorm$g_r)
meltedsimNorm[, b_dist:=as.character(b_dist)]
meltedsimNorm[, sim:=as.character(sim)]
# Alternate versions for controlling plot behavior for b_dist and buffer_field 
meltedsimNorm_2 <- meltedsimNorm
meltedsimNorm_2[, b_dist:=as.factor(b_dist)]
meltedsimNorm_2[, sim:=as.factor(sim)]
fieldNormDF_T_2 <- fieldNormDF_T
fieldNormDF_T_2[, buffer_field:=as.character(buffer_field)]
fieldNormDF_T_2[, buffer_field:=as.factor(buffer_field)]
# Set keys for data.table
setkey(meltedsimNorm, b_dist)

# get 12th largest and smallest values  for each b_dist category (for CI plotting)
y_min <- meltedsimNorm[, .SD[g_r %in% head(sort(unique(g_r)), 13)], by = b_dist]
#y_min_2 <- meltedsimNorm[order(meltedsimNorm$g_r), ]
head(y_min)
#y_min_2[9950:10000, ]
y_minSorted <- y_min[order(b_dist, g_r)]
head(y_minSorted)
#y_minSorted_2 <- y_min_2[order(b_dist, g_r)]

y_minSorted[1:100, ]

y_max <- meltedsimNorm[, .SD[g_r %in% tail(sort(unique(g_r)), 13)], by = b_dist]
y_maxSorted <- y_max[order(b_dist, -g_r)]
y_maxSorted[1:30, ]

# Get every 13th element to set bounds for CI plots 
y_min <- y_minSorted[seq(13, NROW(y_minSorted), by = 13), ]
y_max <- y_maxSorted[seq(13, NROW(y_maxSorted), by = 13), ]
y_lims <- cbind(y_min, y_max)
colnames(y_lims) <- c("b_dist_min", "sim_min", "lwr", "b_dist_max", "sim_max", "upr")

# Calculate summaries of overlap by b_dist group
meltSummary <- summarySE(meltedsimNorm_2, measurevar = "g_r", groupvars = "b_dist")
glimpse(meltSummary)
meltSummary <- cbind(meltSummary, y_lims$lwr, y_lims$upr)
colnames(meltSummary) <- c("b_dist", "N", "g_r", "sd", "se", "ci", "lwr", "upr")


# Test plots of g(r)
ggplot(meltedsimNorm, aes(x=as.numeric(b_dist), y=as.numeric(g_r), group=sim)) + geom_line()

#____________________________________________________________________________
# This is the useful plot with CI's and field data plotted
cols <- c("Simulation mean" = "gray22", "Field data" = "skyblue")
ggplot(meltSummary, aes(x=as.numeric(b_dist), y=g_r)) +
    geom_line(size = 0.1, aes(colour = "Simulation mean")) + 
    geom_ribbon(aes(x = as.numeric(b_dist), ymin = lwr, ymax = upr), alpha = 0.1) +
    geom_line(data = fieldNormDF_T_2, size = 1, aes(as.numeric(buffer_field), g_r_field, colour = "Field data")) + 
    scale_colour_manual(name = "Data Type", values = cols)

# Density plot of overlaps by buffer distance
overlapDens_title <- c("Overlap densities by buffer distance")
overlapDens_plot <- ggplot(meltedsimNorm, aes(g_r, colour = b_dist)) +
    geom_density() +
    ggtitle(overlapDens_title) +
    labs(x = "Proportion overlap", y = 'Density')
overlapDens_plot


# Old code left as reference
# # Correction for polygon approach (Nuske et al. 2009)
# simResults <- bufferDF[1:500, ]
# 
# # Apply normailization to field data NOTE: NEED TO VERIFY THIS IS THE CORRECT WAY TO DO THIS
# field_data <- bufferDF[501, ]
# field_data <- as.data.frame(field_data)
# # Transpose and melt to long format for ggplot2
# simResults_T <- t(simResults)
# head(simResults_T)
# 
# # Coerce simNorm_T & field_Norm_T matrices to dataframes
# simResultsDF_T <- as.data.frame(simResults_T)
# simResultsDF_T <- cbind(simResultsDF_T, bufferDists[2:21])
# colnames(simResultsDF_T) <- c(paste0("sim_", seq(from = 1, to = 500, by = 1), sep = ""), "buffer")
# fieldDataDF <- cbind(field_data, bufferDists[2:21])
# colnames(fieldDataDF) <- c("g_r_field", "buffer_field")
# 
# # Convert to data.table
# setDT(simResultsDF_T)
# meltedsimResults <- melt(simResultsDF_T, id.vars="buffer")
# setDT(fieldDataDF)
# 
# # # Write to .rds **********CHANGE OBJECT AND FILE NAMES***********
# # saveRDS(meltedsimNorm, "./msData/Q20_56_C_meltedSimNorm.rds")
# # saveRDS(fieldNormDF_T, "./msData/Q20_56_C_fieldNormDF_T.rds")
# 
# # Change the types of column data
# glimpse(meltedsimResults)
# names(meltedsimResults) <- c("b_dist", "sim", "g_r")
# meltedsimResults$g_r <- unlist(meltedsimResults$g_r)
# meltedsimResults[, b_dist:=as.character(b_dist)]
# meltedsimResults[, sim:=as.character(sim)]
# glimpse(fieldDataDF)
# fieldDataDF$g_r_field <- unlist(fieldDataDF$g_r_field)
# fieldDataDF[, buffer_field:=as.character(buffer_field)]
# 
# ggplot(meltedsimResults, aes(g_r, colour = b_dist)) +
#     geom_density()

```   




